# GBIF
GBIF (Global Biodiversity Information Facility https://www.gbif.org/) is an open-access biodiversity database. This is an intergovernmental database that started in 2001 and is currently contributed by 97 countries with over 1500 publishers (mainly universities, museums, scientific institutions).

#rgbif
rgbif is an R package to communicate with GBIF, that let you search for information of any taxonomic group of your interest, and count or download their occurrence records.


----------------------------------------------------------------------------------------------
Installation
```{r eval=FALSE, include=FALSE}
install.packages("rgbif")
```
```{r library rgbif}
library(rgbif)
```


----------------------------------------------------------------------------------------------
# Count occurrence records
`occ_count()` will simply let you count occurrence records in GBIF with various arguments.

e.g. 
```{r}
# Count by observation types.
occ_count(basisOfRecord='HUMAN_OBSERVATION')
occ_count(basisOfRecord='MACHINE_OBSERVATION')
occ_count(basisOfRecord='LITERATURE')
occ_count(basisOfRecord='LIVING_SPECIMEN')
occ_count(basisOfRecord='FOSSIL_SPECIMEN')

# Count records that have lat/long data.
occ_count(georeferenced = TRUE)

# Count records in a specific year.
occ_count(year = 2012)

# Count data for a specific taxa.
# Taxa: genus *Ambrosia* (ragweed)
# Taxon key: 3110576
occ_count(taxonKey=3110576)

# Count data for a country.
# Country: United States
# Country code: US
occ_count(country = "US")
```

You can combine multiple data conditions in occ_count arguments.
e.g. Count records of ragweed in USA with lat/long data.
```{r}
# Observation type: Human Observation
# Country: United States (country code = US)
# Taxon: *Ambrosia* (taxonKey= 3110576)
# Has long/lat data
occ_count(basisOfRecord='HUMAN_OBSERVATION', country = "US", taxonKey=3110588, georeferenced = TRUE)
```
Acceptable set of search conditions.
• basisOfRecord, country
• basisOfRecord, country, isGeoreferenced
• basisOfRecord, country, isGeoreferenced, taxonKey
• basisOfRecord, country, taxonKey
• basisOfRecord, datasetKey
• basisOfRecord, datasetKey, isGeoreferenced
• basisOfRecord, datasetKey, isGeoreferenced, taxonKey
• basisOfRecord, datasetKey, taxonKey
• basisOfRecord, isGeoreferenced, taxonKey
• basisOfRecord, isGeoreferenced, publishingCountry
• basisOfRecord, isGeoreferenced, publishingCountry, taxonKey
• basisOfRecord, publishingCountry
• basisOfRecord, publishingCountry, taxonKey
• basisOfRecord, taxonKey
• country
• country, datasetKey, isGeoreferenced
• country, isGeoreferenced
• country, isGeoreferenced, publishingCountry
• country, isGeoreferenced, taxonKey
• country, publishingCountry
• country, taxonKey
• country, typeStatus
• datasetKey
• datasetKey, isGeoreferenced
• datasetKey, isGeoreferenced, taxonKey
• datasetKey, issue
• datasetKey, taxonKey
• datasetKey, typeStatus
• isGeoreferenced
• isGeoreferenced, publishingCountry
• isGeoreferenced, publishingCountry, taxonKey
• isGeoreferenced, taxonKey
• issue
• publishingCountry
• publishingCountry, taxonKey
• publishingCountry, typeStatus
• taxonKey
• taxonKey, typeStatus
• typeStatus
• protocol
• year




# How to get "Taxon key"?
Taxon key can be obtained from a result of `name_backbone()`.
```{r}
name_backbone(name='Ambrosia aptera', rank = "species")
```

e.g. Taxon keys for genus *Ambrosia* and its' species *aptera*.
```{r}
name_backbone(name='Ambrosia', rank = "genus")$genusKey
name_backbone(name='Ambrosia aptera', rank = "species")$speciesKey
```


If you are unsure about taxonomic names and their usages in the GBIF system, you can look them up with `name_suggest()` by putting partial names. 
```{r}
suggest <- name_suggest(q='Ambro')
head(suggest)
```
You can confine output result within specific taxonomic rank.
```{r}
suggest_gen <- name_suggest(q='Ambro', rank = "genus")
head(suggest_gen)
```
If you need more detailed classification of the group, `name_usage()` will return more information.
```{r}
usage <- name_usage(name = "Ambrodiscus", return = "data")
head(usage)
```




# How to get "country code"?
Country codes are stored in "isocodes", a data.frame that comes with rgbif package.
```{r}
head(isocodes)
```
e.g Country code for United States in the GBIF system.
```{r}
isocodes[grep("^United States$", isocodes$name), "code"]
```




# Search for occurrence records
`occ_search()` will return up to 500 occurrence records that are defined in its' argumentation.

occ_search() arguments:
taxonKey = NULL, scientificName = NULL, country = NULL, publishingCountry = NULL, hasCoordinate = NULL, typeStatus = NULL, recordNumber = NULL, lastInterpreted = NULL, continent = NULL, geometry = NULL, geom_big = "asis", geom_size = 40, geom_n = 10, recordedBy = NULL, basisOfRecord = NULL, datasetKey = NULL, eventDate = NULL, catalogNumber = NULL, year = NULL, month = NULL, decimalLatitude = NULL, decimalLongitude = NULL, elevation = NULL, depth = NULL, institutionCode = NULL, collectionCode = NULL, hasGeospatialIssue = NULL, issue = NULL, search = NULL, mediaType = NULL, subgenusKey = NULL, repatriated = NULL, phylumKey = NULL, kingdomKey = NULL, classKey = NULL, orderKey = NULL, familyKey = NULL, genusKey = NULL, establishmentMeans = NULL, protocol = NULL, license = NULL, organismId = NULL, publishingOrg = NULL, stateProvince = NULL, waterBody = NULL, locality = NULL, limit = 500, start = 0, fields = "all", return = "all", spellCheck = NULL, facet = NULL, facetMincount = NULL, facetMultiselect = NULL, skip_validate = TRUE, curlopts = list(),...


e.g.
Taxon: *Ambrosia intergradiens*
Country: Unites States
```{r}
# Country Code
usa_code <- isocodes[grep("^United States$", isocodes$name), "code"]

occurrence_data <- occ_search(scientificName = "Ambrosia intergradiens", country=usa_code, hasCoordinate = TRUE, return='data')
head(occurrence_data)
```




# Obtain occurrence records >500
In order to download large occurrence records at once, you need a GBIF account. Store your user name, password, and email address as your environmental variables (.Rprofile, .Renviron) or simply set them as R objects.
```{r}
user <- "namiko"
pwd <- "735seibutsu"
email <- "nm0076@mix.wvu.edu"
```

1. Make a download request
First, you need to make a download request to the GBIF by using `occ_download()` and let the database compile the data for you. For downloading format, you can choose one from DWCA(Darwin Core Archive), SIMPLE_CSV, and SPECIES_LIST.

e.g.
Taxon: genus *Ambrosia* (key = 3110576)
Country: United States (code = US)
Download format: Darwin Core Archive
```{r}
dl_Ambrosia_USA <- occ_download('taxonKey = 3110576', 'country = US', format = "DWCA", user = user, pwd = pwd, email = email)
dl_Ambrosia_USA
```
Now, you have to wait for a while, from a few minutes up to 3 hours, depending on the amount of data you requested. 
Check the status in https://www.gbif.org/user/download


2. Fetch the data from the database
By using the download key provided by `occ_download()`, you can now download the large occurrence data from the GBIF to your workspace or save them to your computer as a ZIP file. 

e.g. Import data into environment
```{r}
# Download key
dl_Ambrosia_USA <- "0010471-191105090559680"

# occ_download_import(): import data into your workspace
Ambrosia_USA <- occ_download_import(key = dl_Ambrosia_USA)
head(Ambrosia_USA)
```
e.g. Save as a file
```{r}
# occ_download_get(): saves data as a zipfile
occ_download_get(key = dl_Ambrosia_USA, path = "./", overwrite = TRUE,  curlopts = list(verbose=TRUE))
```



# Get citation infomation of the data.
`gbif_citation()` will provide a list of all citations for the datasets that contribute to your occurrence records. The input can be the result of `occ_search()`, `occ_download_get()`, etc.
```{r}
# Country code
usa_code <- isocodes[grep("^United States$", isocodes$name), "code"]
# Occurrence records
occurrence_data <- occ_search(scientificName = "Ambrosia intergradiens", country=usa_code, hasCoordinate = TRUE, return='data')

gbif_citation(occurrence_data)
```




# Browse photos
```{r eval=FALSE, include=FALSE}
media_data <- occ_search(scientificName = "Ambrosia", country = usa_code, mediaType = 'StillImage', return = "media")
gbif_photos(media_data, which='table')
```




# Plot occurrence data on a map
rgbif has a mapping function called `map_fetch()` but it's still under development. Instead, ggplot works well with the occurrence records obtained from GBIF.
```{r}
library(ggplot2)
```

e.g. Ragweed (genus *Ambrosia*) distribution in USA. 
```
# Download data
dl_Ambrosia_USA <- occ_download('taxonKey = 3110576', 'country = US', format = "DWCA", user = user, pwd = pwd, email = email)
Ambrosia_USA <- occ_download_import(key = dl_Ambrosia_USA)
```
```{r}
# Plot with ggplot2
usa <- map_data("usa")
ggplot(usa, aes(long, lat)) +
ggtitle("Species distribution of ragweed (genus Ambrosia)") +
geom_polygon(aes(group = group), fill = "white", 
              color = "black", size = .5) +
geom_jitter(data = Ambrosia_USA, aes(decimalLongitude, decimalLatitude, color = specificEpithet), alpha=1, size = 1, show.legend = TRUE) +
coord_map(xlim = c(-130, -60), ylim = c(20, 50)) +
theme(legend.position="bottom", legend.box = "horizontal", legend.text  = element_text(size = 7), legend.key.size = unit(0.1, "lines"))
```

e.g. Spread of Zika virus, plotted by year. 
```
dl_ZikaVirus <- occ_download('taxonKey = 9848767', format = "DWCA", user = user, pwd = pwd, email = email)
```
```{r}
# Import dowloaded data into environtment
dl_ZikaVirus <- "0007974-191105090559680"
ZikaVirus_world <- occ_download_import(key =dl_ZikaVirus)

# Plot with ggplot2
world <- map_data("world")
ggplot(world, aes(long, lat)) +
ggtitle("Spread of Zika virus -2019") +
geom_polygon(aes(group = group), fill = "white", 
              color = "black", size = .5) +
geom_jitter(data = ZikaVirus_world, aes(decimalLongitude, decimalLatitude, color = year), alpha=1, size = 3, show.legend = TRUE) +
scale_color_gradient(name="Year", low="steelblue1", high="red2")
```